---
title: "Dataset on Suicide Attacks"
author: "Leah Shiferaw"
date: "2024-03-24"
output:
  powerpoint_presentation: default
  beamer_presentation: default
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ggplot2)
library(knitr)
library(readxl)
library(patchwork)


dsat = "dsat_dist_2020_10.xlsx"
df = excel_sheets(dsat) |>
  set_names() |>
  map_dfr(~ read_excel(dsat, sheet = .x), .id = "dsat")

df2 = df |>
  select(date_year, weapon_txt, admin0_txt, group_name)

```

## Overview of the data

The Database on Suicide Attacks (DSAT) was created by researchers at the 
University of Chicago's Project on Security and Threats (CPOST) and 
examines worldwide suicide attacks from 1982 to 2019. The data were collected through database searches, Google alerts, search engine and social media searches.


The goal of this analysis is to understand what trends exist in the weapons, countries, years, and groups that conducted suicide attacks. 



## Overview of variables

Top groups conducting suicide attacks

```{r top groups}

top_groups = df2 |>
  filter(!is.na(group_name)) |>
  count(group_name, name = "attack_count") |>
  arrange(desc(attack_count)) |>
  head(5)

kable(top_groups, caption = "Top Five Groups")
knitr::kable(top_groups, caption = "Top Five Groups")

```

## Top countries in which suicide attacks occurred

```{r top countries}

top_countries = df2 |>
  filter(!is.na(admin0_txt)) |>
  count(admin0_txt, name = "attack_location") |>
  arrange(desc(attack_location)) |>
  head(5)

kable(top_countries, caption = "Top Five Countries")
knitr::kable(top_countries, caption = "Top Five Countries")

```

## Analysis 1: Suicide attacks by year


```{r}

summary(df2$date_year)

df3 = df2 |>
  count(date_year) |>
  rename(n = n)

top_years = df2 |>
  filter(!is.na(date_year)) |>
  count(date_year, name = "attack_year") |>
  arrange(desc(attack_year)) |>
  head(5)

kable(top_years, caption = "Top Five Attack Years")
knitr::kable(top_years, caption = "Top Five Attack Years")

ggplot(df3, aes(x = date_year, y = n)) +
  geom_line(color = "#0069c0") +
  geom_point(color = "black", size = 1) +
  labs(x = "Year", y = "Count", title = "Suicide Attacks by Year") +
  scale_y_continuous(limits = c(0, 1000)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

date1 = subset(df3, date_year >= 1982 & date_year <= 2000)

plot1 = ggplot(date1, aes(x = date_year, y = n)) +
  geom_line(color = "#0069c0") +
  geom_point(color = "black", size = 2) +
  labs(x = "Year", y = "Count", title = "Suicide Attacks: 1982 to 2000") +
  scale_y_continuous(limits = c(0, 1000)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

date2 = subset(df3, date_year >= 2001 & date_year <= 2019)

plot2 = ggplot(date2, aes(x = date_year, y = n)) +
  geom_line(color = "#0069c0") +
  geom_point(color = "black", size = 2) +
  labs(x = "Year", y = "Count", title = "Suicide Attacks: 2001 to 2019") +
  scale_y_continuous(limits = c(0, 1000)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
print(plot1)
print(plot2)

combined_plots <- plot1 + plot2
combined_plots
print(combined_plots)

```

## Analysis 2: Most common types of weapons

The researchers have identified 17 types of weapons used to carry out suicide attacks. The top five most frequently used weapons were:

```{r weapon types}

weapons = df2 |>
  count(weapon_txt) |>
  arrange(desc(n)) |> 
  filter(!is.na(weapon_txt))
  
top_weapons = head(weapons, 5)  

kable(top_weapons, caption = "Top Five Weapons")
knitr::kable(top_weapons, caption = "Top Five Weapons")

```

## Weapon Types by Year


```{r top 5 weapons}

weapon_counts = df2 |>
  count(date_year, weapon_txt) |>
  arrange(date_year, desc(n)) |>
  group_by(date_year) |>
  slice_head(n = 5) |>
  ungroup()

selected_weapons = c("Car bomb", "Belt bomb", "Truck bomb", "Unspecified", "Motorcycle bomb")


filtered_df = df2 |>
  filter(weapon_txt %in% selected_weapons)

weapon_counts = filtered_df |>
  count(date_year, weapon_txt) |>
  arrange(date_year, desc(n)) |>
  group_by(date_year) |>
  slice_head(n = 5) |>
  ungroup()

```

```{r combined plots, fig.width=11, fig.height=6}
plot3 = ggplot(top_weapons, aes(x = reorder(weapon_txt, n), y = n, fill = weapon_txt)) +
  geom_bar(stat = "identity") +
  labs(x = "Weapon Type", y = "Count", title = "Top Five Most Frequent Weapon Types") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

plot4 = ggplot(weapon_counts, aes(x = date_year, y = n, color = weapon_txt)) +
  geom_line() +
  labs(x = "Year", y = "Count", title = "Top Five Weapons by Year") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

combined_plots <- plot3 + plot4
combined_plots

print(ggplot)


```



